
# I.Setup
Libraries needed: <!-- Note: I wrote results hide but the ## are not results maybe?? -->[Note: I wrote results hide but the ## are not results maybe??]
```{r, results = "hide"}
library(Biobase)
library(GEOquery)
library(Seurat)
library(CellChat)
library(dplyr)
library(patchwork)
```
```{r}
load(url("https://ndownloader.figshare.com/files/25950872"))

data.input = data_humanSkin$data
meta = data_humanSkin$meta
husk <- CreateSeuratObject(data.input, meta.data = meta)

husk
husk.data
husk.data[c("A1BG", "A4GALT", "A4GALT"), 1:30]
```



# II. Pre-processing

```{r}
husk[["percent.mt"]] <- PercentageFeatureSet(husk, pattern = "^MT-") 
VlnPlot(husk, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3) 
head(husk@meta.data, 5) 
head(husk@meta.data, 10)

husk.data[c("CCL2", "CCL19", "CCR7"), 1:30] 
husk.data[c("CCL2", "CCL19", "CCR7"), 2000:2020]
plot1 <- FeatureScatter(husk, feature1 = "nCount_RNA", feature2 = "percent.mt") 
plot2 <- FeatureScatter(husk, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") 
plot1 + plot2
```
filter "husk"
```{r}
husk <- subset(husk, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)
```


### Visualization of feature-feature relationships
```{r}
plot1 <- FeatureScatter(mouse, feature1 = "nCount_RNA", feature2 = "percent.mt") 

#since all have same mt quality...

plot2 <- FeatureScatter(mouse, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
#positively correlated

plot1 + plot2
```

# III. Normalization
husk <- NormalizeData(husk)

# IV. Feature Selection
```{r}
husk <- FindVariableFeatures(husk, selection.method = "vst", nfeatures = 2000)
top10 <- head(VariableFeatures(husk), 10) 
top10

top20 <- head(VariableFeatures(husk), 20) 
top20 

top50 <- head(VariableFeatures(husk), 50) 
top50

plot1 <- VariableFeaturePlot(husk)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```

# V. Scaling
```{r}
all.genes <- rownames(husk)
husk <- ScaleData(husk, features = all.genes)
```
# VI. Linear Dimension Reduction
### Perform PCA
```{r}
husk <- RunPCA(husk, features = VariableFeatures(object = husk))
print(husk[["pca"]], dims = 1:5, nfeatures = 5)
VizDimLoadings(husk, dims = 1:2, reduction = "pca")
DimPlot(husk, reduction = "pca")
DimPlot(husk, reduction = "pca",label = TRUE)
```

heatmap: exploration of the primary sources of heterogeneity -> decide which PCs to include for further downstream analyses
```{r}
DimHeatmap(husk, dims = 1, cells = 500, balanced = TRUE) #heatmap: exploration of the primary sources of heterogeneity
DimHeatmap(husk, dims = 1:15, cells = 500, balanced = TRUE)
```
# VII. Determination of Dimension
Metafeature combines information across a correlated feature set. 
top principal components => robust compression
how many PCs to take?
###JackStraw procedure
permute subset of the data (1% by default) and rerun PCA
significant PCs ~ strong enrichment of low p-value features
```{r}
husk <- JackStraw(husk, num.replicate = 100)
husk <- ScoreJackStraw(husk, dims = 1:20)
```
```{r}
JackStrawPlot(husk, dims = 1:15)
```

alternative heuristic method Elbow plot:
observe where the elbow is and then see which PCs to choose
```{r}
ElbowPlot(husk)
```
# VIII. Clustering Cells
```{r}
husk <- FindNeighbors(husk, dims = 1:10)
husk <- FindClusters(husk, resolution = 0.7)
head(Idents(mouse), 5)
```
# IX. Non-linear Dimension Reduction
## e.g. UMAP, tSNE
input to the UMAP and tSNE: same PCs as input to the clustering analysis
```{r}
husk <- RunUMAP(husk, dims = 1:10)
#DimPlot(husk, reduction = "umap")
DimPlot(husk, reduction = "umap", label = TRUE)
saveRDS(husk, file = "/Users/xiaoyizheng/Downloads/Bioinfo_Einarbeiten/CellChatSelf/husk/output/huskOutput.rds")
```


# X. Cluster Biomarkers
### Finding differentially expressed features
- min.pct argument:  requires a feature to be detected at a minimum percentage in either of the two groups of cells
- thresh.test argument: requires a feature to be differentially expressed (on average) by some amount between the two groups
- to speed up computations: max.cells.per.ident
```{r}
cluster0.markers <- FindMarkers(husk, ident.1 = 0, min.pct = 0.25)
head(cluster0.markers, n = 5)
VlnPlot(husk, features = c("CLEC3B", "FBN1", "FSTL1", "IGFBP5", "MFAP5"))

cluster1.markers <- FindMarkers(husk, ident.1 = 1, min.pct = 0.25)
head(cluster1.markers, n = 5)
VlnPlot(husk, features = c("APOE", "CXCL12", "IGFBP7", "APOC1", "MGST1"))

cluster2.markers <- FindMarkers(husk, ident.1 = 2, min.pct = 0.25)
head(cluster2.markers, n = 5)
VlnPlot(husk, features = c("B2M", "CD3D", "CD52", "GSN", "IL32"))


cluster3.markers <- FindMarkers(husk, ident.1 = 3, min.pct = 0.25)
head(cluster3.markers, n = 5)
VlnPlot(husk, features = c("CFD", "APOD", "GSN", "DCN", "CXCL14"))

cluster4.markers <- FindMarkers(husk, ident.1 = 4, min.pct = 0.25)
head(cluster4.markers, n = 5)
VlnPlot(husk, features = c("CD69", "FOS", "JUNB", "CD52", "DNAJB1"))

cluster5.markers <- FindMarkers(husk, ident.1 = 5, ident.2 = c(0, 3), min.pct = 0.25)
head(cluster5.markers, n = 5)
VlnPlot(husk, features = c("COL6A5", "TNC", "APCDD1", "COL18A1", "CFD"))

husk.markers <- FindAllMarkers(husk, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.2) #0.2 is okki
husk.markers %>%
  group_by(cluster) %>%
  slice_max(n = 2, order_by = avg_log2FC)

FeaturePlot(husk, features = c("CFD", "IBSP", "COCH", "CCL1", "HLA-DRA", "APOD", "MMP12", "CXCL14", "CCL19", "CD74"))

husk.markers %>%
  group_by(cluster) %>%
  top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(husk, features = top10$gene) + NoLegend()

```

# XI. Assigning Cell Type Identity to Clusters
```{r}
new.cluster.ids <- c("0","1","FIB","3","TC","5","6","7","8","9","10")
names(new.cluster.ids) <- levels(husk)
husk <- RenameIdents(husk, new.cluster.ids)
DimPlot(husk, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
```
