---
title: "md"
output: html_document
date: "2022-12-01"
---
# I.Setup
Libraries needed: <!-- Note: I wrote results hide but the ## are not results maybe?? -->[Note: I wrote results hide but the ## are not results maybe??]
```{r, results = "hide"}
library(dplyr)
library(Seurat)
library(patchwork)
```

Loading data & Checking out how the loaded data looks like:
```{r, results = "hide"}
mouse.data <- Read10X(data.dir = "/Users/xiaoyizheng/Downloads/Praktikum/Bioinfo_Einarbeiten/CellChatSelf/mouse/GSE113854_RAW/")

mouse.data
```

Creating Seurat object:
```{r}
mouse <- CreateSeuratObject(counts = mouse.data, project = "MoSk", min.cells = 3, min.features = 200)
```
# II. Pre-processing
QC (quality control) Matrix for selection and filtration of cells, using user-defined criteria. In this case it is the mitochondrial genes as criteria. [note: apparently they all have same level of mt content. Maybe pre-processed already? or maybe that is not the proper criteria to choose in this case. Which one could be better instead of it?]
```{r}
mouse[["percent.mt"]] <- PercentageFeatureSet(mouse, pattern = "^MT-") 

#Visualization by violin plot (mt percent no significance):
VlnPlot(mouse, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3) 
```
Genes and total molecules calculated and stored in obj. meta data.
Showing QC metrics for the first 5 cells:
```{r}
head(mouse@meta.data, 5) 
```

```{r}
#examining a few genes in the first thirty cells
mouse.data[c("Igf2", "Cck", "Pf4"), 1:30] 
#a few genes in the first 30 cells 
#picked "Igf2", "Cck", "Pf4" according to web
```


### Visualization of feature-feature relationships
```{r}
plot1 <- FeatureScatter(mouse, feature1 = "nCount_RNA", feature2 = "percent.mt") 

#since all have same mt quality...

plot2 <- FeatureScatter(mouse, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
#positively correlated

plot1 + plot2
```

# III. Normalization
since it is already normalized => this step omitted

# IV. Feature Selection
```{r}
mouse <- FindVariableFeatures(mouse, selection.method = "vst", nfeatures = 2000)
top10 <- head(VariableFeatures(mouse), 10) # Identify the 10 most highly variable genes
top10

plot1 <- VariableFeaturePlot(mouse)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```

# V. Scaling
```{r}
all.genes <- rownames(mouse)
mouse <- ScaleData(mouse, features = all.genes)


#optional: mouse <- ScaleData(mouse)
```
# VI. Linear Dimension Reduction
### Perform PCA
```{r}
mouse <- RunPCA(mouse, features = VariableFeatures(object = mouse))
```
### Examine and visualize PCA results a few different ways
```{r}
print(mouse[["pca"]], dims = 1:5, nfeatures = 5)
```
```{r}
VizDimLoadings(mouse, dims = 1:2, reduction = "pca")
```
```{r}
DimPlot(mouse, reduction = "pca")
```

heatmap: exploration of the primary sources of heterogeneity -> decide which PCs to include for further downstream analyses
```{r}
DimHeatmap(mouse, dims = 1, cells = 500, balanced = TRUE) #heatmap: exploration of the primary sources of heterogeneity
DimHeatmap(mouse, dims = 1:15, cells = 500, balanced = TRUE)
```
# VII. Determination of Dimension
Metafeature combines information across a correlated feature set. 
top principal components => robust compression
how many PCs to take?
###JackStraw procedure
permute subset of the data (1% by default) and rerun PCA
significant PCs ~ strong enrichment of low p-value features
```{r}
mouse <- JackStraw(mouse, num.replicate = 100)
mouse <- ScoreJackStraw(mouse, dims = 1:20)
```
```{r}
JackStrawPlot(mouse, dims = 1:15)
```

alternative heuristic method Elbow plot:
observe where the elbow is and then see which PCs to choose
```{r}
ElbowPlot(mouse)
```
# VIII. Clustering Cells
```{r}
mouse <- FindNeighbors(mouse, dims = 1:10)
mouse <- FindClusters(mouse, resolution = 0.5)

# Look at cluster IDs of the first 5 cells
head(Idents(mouse), 5)
```
# IX. Non-linear Dimension Reduction
## e.g. UMAP, tSNE
input to the UMAP and tSNE: same PCs as input to the clustering analysis
```{r}
mouse <- RunUMAP(mouse, dims = 1:10)
#note that you can set `label = TRUE` or use the LabelClusters function to help label individual clusters
DimPlot(mouse, reduction = "umap")
#saveRDS(mouse, file = "/Users/xiaoyizheng/Downloads/Bioinfo_Einarbeiten/CellChatSelf/mouse/output/mouseOutput.rds")
```


# X. Cluster Biomarkers
### Finding differentially expressed features
- min.pct argument:  requires a feature to be detected at a minimum percentage in either of the two groups of cells
- thresh.test argument: requires a feature to be differentially expressed (on average) by some amount between the two groups
- to speed up computations: max.cells.per.ident
```{r}
# find all markers of cluster 2
cluster2.markers <- FindMarkers(mouse, ident.1 = 2, min.pct = 0.25)
head(cluster2.markers, n = 5)
cluster5.markers <- FindMarkers(mouse, ident.1 = 5, ident.2 = c(0, 3), min.pct = 0.25)
head(cluster5.markers, n = 5)
mouse.markers <- FindAllMarkers(mouse, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)  #No features pass logfc.threshold threshold, maybe threshold=0.1, 0.15, 0.2?
mouse.markers <- FindAllMarkers(mouse, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.2) #0.2 is okki
mouse.markers %>%
  group_by(cluster) %>%
  slice_max(n = 2, order_by = avg_log2FC)
#cluster0.markers <- FindMarkers(mouse, ident.1 = 0, logfc.threshold = 0.25, test.use = "roc", only.pos = TRUE)
cluster0.markers <- FindMarkers(mouse, ident.1 = 0, logfc.threshold = 0.2, test.use = "roc", only.pos = TRUE)
VlnPlot(mouse, features = c("Rgs5", "Tyrobp"))
VlnPlot(mouse, features = c("Col1a1", "Col1a2"), slot = "counts", log = TRUE) #plot raw counts
FeaturePlot(mouse, features = c("Sparc", "Tyrobp", "Igfbp7", "Pecam1", "Cd34", "Mfap4", "Ndufa4l2", "Il6",
                               "Birc5"))
mouse.markers %>%
  group_by(cluster) %>%
  top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(mouse, features = top10$gene) + NoLegend()
```

# XI. Assigning Cell Type Identity to Clusters
new.cluster.ids <- c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15")
names(new.cluster.ids) <- levels(mouse)
mouse <- RenameIdents(mouse, new.cluster.ids)
DimPlot(mouse, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()

#FindConservedMarkers(mouse, idents.1= [nr.] )
#Idents(mouse)
#RenameIdents(mouse, '[nr.]'= [celltype] )